Simultaneous segmentation and distortion correction on diffusion weighted MR using shape priors

### 2012-12-21 Publications/2013-IPMI/Latex/03-Introduction.tex

We suggest clustering the current methodologies of template-based segmentation 
methods into three groups. The first group typically adds a shape prior term to 
the energy functional of an evolving active contour~\citep{Rousson2002,Chen2002,
Paragios2003,Vemuri2003a,Yezzi2003a,Gastaud2004,Chan2005,Cremers2006,Bresson2006a,
Ayvaci2007,Schmid2008} 
{\color{red}{(FIXME:I would select from here the most important,
removing directly those appearing in conferences and never published on journal. Better
if we found a review paper/book chapter. I don't like so many references on a row)}}.
These methods generally have a explicit description of the expected relative boundary 
locations of the object to be delineated, and some even model the statistical deviations
from this average shape. Closely related to this group are atlas-based segmentation
methods \citep{Pohl2005,Pohl2006,Wang2006,Gorthi2009,Gorthi2011}, where the prior 
imposes consistent voxel-based classification of contiguous regions. Here, the 
presence of more structures than one unique \ac{roi} helps aligning the target image 
with the atlas in a hierarchical fashion. Finally, the third group generalizes 
the atlas to actual images, and the contour is to segment simultaneously two 
different target images, related by a spatial transform to be co-estimated~\citep{Wyatt2003,
Yezzi2003}.

>>>>

We suggest clustering the current methodologies of template-based segmentation 
methods into three groups. The first group typically adds a shape prior term to 
the energy functional of an evolving active contour~\citep{Chen2002,
Paragios2003,Vemuri2003a,Yezzi2003a,Gastaud2004,Chan2005,Cremers2006,Bresson2006a} %removed Rousson2002,Ayvaci2007,Schmid2008
%{\color{red}{(FIXME:I would select from here the most important,
%removing directly those appearing in conferences and never published on journal. Better
%if we found a review paper/book chapter. I don't like so many references on a row)}}.
These methods generally have a explicit description of the expected relative boundary 
locations of the object to be delineated, and some even model the statistical deviations
from this average shape. Closely related to this group are atlas-based segmentation
methods \citep{Pohl2005,Pohl2006,Wang2006,Gorthi2009,Gorthi2011}, where the prior 
imposes consistent voxel-based classification of contiguous regions. Here, the 
presence of more structures than one unique \ac{roi} helps aligning the target image 
with the atlas in a hierarchical fashion. Finally, the third group generalizes 
the atlas to actual images, and the contour is to segment simultaneously two 
different target images, related by a spatial transform to be co-estimated~\citep{Wyatt2003,
Yezzi2003}.

### 2012-12-21 Publications/2013-IPMI/Latex/05-Results.tex

\subsection{Synthetic gray-scale data}
The proposed method solved precisely the specific challenge
created by the model. A severe distortion of the model is
artificially created adding complexity to the problem of
partial voluming in the outer contour of the sulcus.
\autoref{fig:sulcusmodel_result} provides visual assessment
for this result. With 16x16x16 control points, 
computation time for this model was around
10 minutes in a {\color{red}{(specific machine details)}}.
\begin{figure}
\includegraphics[width=1.0\textwidth]{model2result}
\caption{Result of the segmentation performance on the first
model. In red and blue colors the obtained \ac{wm}/\ac{gm} 
and \ac{gm}/\ac{csf} contours for the deformed image, respectively.
In white color, the initial shape priors for the \ac{wm}/\ac{gm}
interface.
{\color{red}{This figure is to be replaced by the real result,
this is a proof of concept of registration that does not solve
the pv problem. The current figure also removes the outer prior}}}
\label{fig:sulcusmodel_result}
\end{figure}

>>>>

\subsection{Synthetic gray-scale data}
The proposed method solved precisely the specific challenge
created by the model. A severe distortion of the model is
artificially created adding complexity to the problem of
partial voluming in the outer contour of the sulcus.
\autoref{fig:sulcusmodel_result} provides visual assessment
for this result. With 16x16x16 control points, 
computation time for this model was around
10 minutes in a {\color{red}{(specific machine details)}}.

\begin{figure}
\begin{tabular}{ccccc}
\includegraphics[width=0.19\textwidth]{model2result_b_1} &
\includegraphics[width=0.19\textwidth]{model2result_b_2} &
\includegraphics[width=0.19\textwidth]{model2result_a_1} &
\includegraphics[width=0.19\textwidth]{model2result_a_2} &
\includegraphics[width=0.19\textwidth]{model2surf} \\
\multicolumn{2}{c}{a)} & \multicolumn{2}{c}{b)} & c)
\end{tabular}
\caption{Result of the segmentation performance on the sulcus model.
a) Two slices and prior contours before segmentation-registration. b) The same slices and contours after distortion estimation. c) 3D rendering of the deformed priors, with the traces of the two selected slices highlighted. 
%In red and blue colors the obtained \ac{wm}/\ac{gm} 
%and \ac{gm}/\ac{csf} contours for the deformed image, respectively.
%In white color, the initial shape priors for the \ac{wm}/\ac{gm}
%interface.
{\color{red}{This figure is to be replaced by the real result,
this is a proof of concept of registration that does not solve
the pv problem. The current figure also removes the outer prior}}}
\label{fig:sulcusmodel_result}
\end{figure}
